
\addcontentsline{toc}{chapter}{Abstract}
\begin{abstract}
	Contemporary infantry warfare is defined by the transformation of exposure into the central organising variable of combat: who is seen, when they are seen and how quickly observation becomes fire. Drawing on scholarly and doctrinal sources, practitioner interviews and thematic analysis of open-source combat evidence from the Russo--Ukrainian War, this thesis examines how uncrewed systems reshape light-infantry tactics, mission command and small state force design. The democratisation of ISR and precision strike reconceptualises movement as timed exposure across space, time and the electromagnetic spectrum. Persistent surveillance imposes an infantry ceiling, where dispersion preserves survivability but precludes the concentration of mass required for breakthrough. Manoeuvre becomes conditional, achieved through the sequencing of ISR, EW and fires. Tactically, these dynamics drive section-level adaptation: the 14-soldier section emerges as the drone-survival minimum, integrating organic ISR, CUAS, assault pioneering and additional junior leadership. Strategically, for small states such as Ireland, low-cost uncrewed systems offer a means of supporting deterrence-by-denial through disproportionate cost imposition, provided they are integrated into doctrine, training, command, sustainment and force design.
\end{abstract}

\cleardoublepage

\subsection*{Summary of Analytical Contributions}
\addcontentsline{toc}{chapter}{Summary of Analytical Contributions}
\small
\renewcommand{\arraystretch}{1.1}
\setlength{\tabcolsep}{4pt}

\begin{tabularx}{\textwidth}{@{} >{\bfseries}l X @{}}
	\toprule
	\multicolumn{1}{c}{\textbf{Concept}} & \multicolumn{1}{c}{\textbf{Analytical Contribution}} \\
	\midrule
	
	Movement as Timed Exposure 
	& Reconceptualises movement as the management of exposure across space, time and the electromagnetic spectrum. \\
	
	Infantry Ceiling 
	& Identifies the threshold at which dispersion precludes the concentration of mass required for breakthrough. \\
	
	Conditional Manoeuvre 
	& Reframes manoeuvre as a phase enabled by the sequencing of ISR, EW and fires rather than a primary generator of success. \\
	
	Least Interesting Target 
	& Defines survivability as minimising signature and target value within the enemy's targeting calculus. \\
	
	FPVD as Organising Weapon 
	& Positions armed drones as an emerging organising weapon around which assault elements may be structured. \\
	
	Section-Level Adaptation 
	& Demonstrates the requirement for organic ISR, CUAS, IEDD and expanded junior leadership at section level. \\
	
	Vertical IED Hazard 
	& Extends the IED threat vertically, imposing a render-safe burden and driving the need for organic IEDD. \\
	
	Mission Command Reinforced 
	& Shows that persistent ISR increases the necessity of decentralised decision-making under degraded control. \\
	
	UxS and Asymmetric Deterrence 
	& Establishes that low-cost UxS enable disproportionate combat power through cost imposition. \\
	
	Drone Shock
	& Identifies the psychological and tactical disruption caused by persistent drone presence, where the sound or expectation of UAS attack alters movement, attention and immediate action. \\
	
	Relative Combat Power
	& Demonstrates that UxS impose a disproportionate cost on attacking forces. \\
	
	\bottomrule
\end{tabularx}
