\addchap*{Acknowledgements} 
\addcontentsline{toc}{chapter}{Acknowledgements}

To Jeremy Maxwell: Your calm supervisory style was greatly appreciated. I always felt that I was heading in the right direction --- even if I changed tracks a few times. I suspect my word count was a recurring source of concern, but your patience and steady guidance were invaluable.

To Brian \textit{Henry} McIntyre: As noted in person, commuting together proved exceptionally important for my mental health during the course. I suspect you are sick of hearing about drones by now.

To Eoin O'Brien: I didn't want to live two courses, yet the reverse was true! Thanks for your patience and willingness to provide sound advice.

To George O'Connell: This thesis began germinating at the \textit{Future Soldier Technology Conferences} in 2023 and 2024, after you invited me to attend. I have grown as an Ordnance technical officer over the past few years --- largely by taking my cue from you --- and that growth is reflected throughout this work.

To Fergal MacDonald: \textit{Robot dogs}, again! You quietly gave me support --- and guardrails --- for my enthusiasm for this topic as it developed in my mind while in J3/5. Your humour helped me keep perspective and your encouragement gave me confidence that the idea was worth pursuing.

To Aisling: The JCSC took more out of me than I had said it would. Knowing me --- in many ways better than I know myself --- you saw through the rhetoric: Chac m\'e ar na huibheacha. Without your support, this thesis would not be what it is. Work has once again intruded on home life and placed an undue burden at your feet. \textit{My} thesis is appearing in \textit{your} nightmares! We welcomed Ois\'{i}n in December and you've taken it all in your stride. I've been humbled. This work is therefore dedicated to you --- \textit{my biggest advocate}.
\linebreak
\textbf{AB, June 2026}

